---
title: "WUP_04"
author: "Eduardo Noriega"
date: "Fall 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
data = read.csv ('GSScsv.csv')
```



# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE 

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We have to predict WORDSUM based on degree.   
WORDSUM is numerical, and degree is categorical. So I used the numerical, categorical helper file.  
 
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
par(mar = c(5, 15, 5, 2))

boxplot(WORDSUM~DEGREE,data=data)
```
 
 On the box plot, it shows that there is a difference on WORDSUM, based on the degree. High school and less than high school degrees have the lowest WORDSUM. While those with a graduate degree has the highest, and the bachelor's has the second highest.  

### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=data)
```
  
 For the numerical summary, this table confirms that the graduate's median is 11, indicating the highest in WORDSUM. Also, the lowest WORDSUM median is 8 coming from the less than (LT) High School, scoring the lowest, and high school graduates scoring the second lowest.   
 
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
model1 <- aov(WORDSUM~DEGREE,data=data)
summary.aov(model1)
```
The null hypothesis is rejected due to p-value being less than 0.005. The p-value is 0.000236, which means that there is at least one difference in WORDSUM based on the aov model.

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 17, 2, 2))
plot(TukeyHSD(model1),las=2)
```
  
  In addition, the TukeyHSD model also confirms that the differences are in graduates, less than high school, and high school graduates.
  

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The null hypothesis is rejected, because there is a difference in WORDSUM scores based on the level of degree. As shown in the numerical summary the mean WORDSUM scores for graduates was 10.09, bachelors means was 9.20, high school 8.28, and less than high school scored 7.75. The TUKEY HSD model as previously stated, shows a difference less than high and graduate degrees; high school and graduates also show a differnce in WORDSUM based on the level of school.
